---
title: "Sample Script"
subtitle: "Parameter Estimation Demo"
author: "Marie Bouffard"
date: '2022-05-23'
output:
  html_document:
    theme: cosmo
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# Attach Packages
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(boot)
library(gt)
library(patchwork)
library(broom)
library(nlstools)

# Read in data and wrangle
fish_catch <- read.csv(here("Data", "fish_catch.csv"), col.names = c("year", "wild_catch", "farmed_fish", "total_fish_production")) %>% 
  mutate(year = as.numeric(year)) %>%
  mutate(wild_catch = as.numeric(wild_catch)) %>% 
  slice(5:67)

# Count up from year 0 at 1950
catch_year <- fish_catch %>% 
  mutate(start_year = year - min(year))

```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

The following sample code comes from part of an assignment I completed for my Advanced Environmental Data Analysis course at the Bren School of Environmental Science and Management on 3/4/2021.

</div>

### Initial Parameter estimates

**a.)** Exploratory plot of wild catch over time.

```{r}

# Exploratory plot

ggplot(data = catch_year, aes(x = year, y = wild_catch)) +
  geom_point(color = "aquamarine3") +
  labs(
    title = "Wild catch exploratory graph",
    x = "Year",
    y = "Global wild fish catch (million tons)"
  ) +
  theme_minimal()

```

**b.)** This trend looks like it is following logistic growth where $K$ is about 90 million tons and $P_o$ appears to be around 17 million tons.

We can estimate $A$ at about 4.3 million tons using the equation $A=\frac{K-P_o}{P_o}$

We can estimate the growth rate constant during the exponential phase by fitting a linear model and using the slope to estimate $k$.

```{r}

# filter for exponential section betweetn 1950 and 1990
wild_catch_exp <- catch_year %>%
  select(start_year, wild_catch) %>% 
  filter(start_year < 40) %>%
  mutate(ln_catch = log(wild_catch))

# Linear model
lm_k <- lm(ln_catch ~ start_year, data = wild_catch_exp)
lm_k

```

This gives us our initial estimate for $k$, 0.035, along with our estimates for $K$ ~ 90 and $A$ ~ 4.3.

### Nonlinear Least Squares

```{r}

# Model info with estimated starting parameter values
wild_nls <- nls(wild_catch ~ K/(1 + A*exp(-r*start_year)),
              data = catch_year,
              start = list(K = 90, A = 4.3, r = 0.035),
              trace = TRUE)

summary(wild_nls)

```

**c.)** Parameter outcomes

```{r}

wild_out_tidy <- broom::tidy(wild_nls)

wild_out_tidy

```

Wild catch model with estimated parameters:
$$P(t) = \frac{100.27}{1+4.31e^{0.069t}}$$

**d.)** Visualization of original data and model output

```{r}

# catch predictions
catch_predict <- predict(wild_nls)

# combine predictions and original data
wild_complete <- data.frame(catch_year, catch_predict)

# Plot together
ggplot(data = wild_complete, aes(x = year, y = wild_catch)) +
  geom_point(color = "aquamarine3") +
  geom_line(aes(x = year, y = catch_predict), color = "aquamarine4") +
    labs(
    title = "Global Wild Fish Catch Model and Original Data",
    x = "Year",
    y = "Catch (million tons)"
  ) +
  theme_minimal()

```

***

**Data Source**
Global wild fish catch and aquaculture production, compiled by Earth Policy Institute with 1950-2010 from U.N. Food and Agriculture Organization (FAO), Global Capture Production and Global Aquaculture Production, electronic databases, at www.fao.org/fishery/topic/16140/en.